#Part one: Introduction to data

This is a analysis of EDA Medical Apppointment No-show data. Data is from [Kaggle](https://www.kaggle.com/joniarroba/noshowappointments). First we are going to have a good overview of the data and then we are going to predict show-ups and no-shows with some statistical modelling.

This analysis is the final assignment of the IODS course. You can find the [GitHub repository here.](https://github.com/mkkjrv/IODS-final) Author of this analysis is Mikko Järvi *mikko.jarvi(at)helsinki.fi*

```{r, include = FALSE}
library(ggplot2)
library(GGally)
library(lubridate)
library(tidyverse)
library(gridExtra)
```

##Overview of data##

After cleaning the the original dataset there is 15 variables and 300 000 observations. All the variables were characters in the original dataset. For the analysis the variables were converted as follows. Script for the data wrangling is [here.](https://github.com/mkkjrv/IODS-final/blob/master/data/create_noshow.R)

```{r, echo = FALSE}
noshow <- read.table("data/noshow.txt", header = TRUE, sep = "\t")
noshow$DayOfTheWeek <- factor(noshow$DayOfTheWeek,
        levels = c("Monday", "Tuesday", "Wednesday", "Thursday", "Friday" , "Saturday", "Sunday"))
noshow$AppointmentRegistration <- date(noshow$AppointmentRegistration)
noshow$AppointmentData <- date(noshow$AppointmentData)
noshow$RegYear <- as.factor(noshow$RegYear)
noshow$RegMonth <- as.factor(noshow$RegMonth)
noshow$RegDay <- as.factor(noshow$RegDay)
noshow$AppYear <- as.factor(noshow$AppYear)
noshow$AppMonth <- as.factor(noshow$AppMonth)
noshow$AppDay <- as.factor(noshow$AppDay)
str(noshow)
```

When you look at the variables you could easily divide them for three groups. First group includes basic information about the patient like *age* and *gender*. Second group includes information about the diseases and habits of the patient like *diabetes*, *alcoholism* and wheter patient smokes or not. Third group of variables includes infromation about the time of the registration and the appointment.

###Research question###

In this analysis I'm going to find the factors that influences most on patients no-shows.I'm also going to create statistical model to predict when patient is most likely going to miss the appointment. 


###Status###

```{r, echo = FALSE}
ggplot(noshow, aes(x=Status, fill = Status)) + geom_bar()
``` 

```{r, echo = FALSE }
status_table <- table(noshow$Status)
status_table
```

Propotion of the patients that don't show up is high. Seems that the dataset is not random.

###Sex and age###
```{r}
#a boxplot of high_use, sex and age
g1 <- ggplot(noshow, aes(x = Status, y = Age, col = Gender))
g1 + geom_boxplot() + ggtitle("Patients by age and sex") + xlab("No-show and show up groups") + ylab("Age of patients")
```

It seems that the distribution of age is similar in both patient groups and there is no significant difference between males and females.

##No-Shows in time##

```{r, echo = FALSE}
table_AppYear <- table(noshow$Status, noshow$AppYear)
addmargins(table_AppYear)

ggplot(noshow, aes(x=AppYear, fill=Status )) + geom_bar(position="fill") + xlab("Propotions of patients by year")
```

Propotion of the patient gropus seems to be identical between both years.


```{r, echo = FALSE}
table_AppMonth <- table(noshow$Status, noshow$AppMonth)
addmargins(table_AppMonth)

ggplot(noshow, aes(x=AppMonth, fill=Status )) + geom_bar(position="fill") + xlab("Propotions of patients by month")
```

The propotion of patients who do not show up on appointment is steady during months. There seems to be slight trend that at end of the year no-shows are increasing but the effect seesm to be modest.


```{r, echo = FALSE}
table_DayOfTheWeek <- table(noshow$Status, noshow$DayOfTheWeek)
addmargins(table_DayOfTheWeek)

ggplot(noshow, aes(x=DayOfTheWeek, fill=Status )) + geom_bar( position = "fill") + xlab("Propotions of patients by weekday")
```

It seems that there is no significant difference in patient groups by weekday. At sundays there seems to be less no-shows but on the other hand there is only 6 patients in two years on sundays.

```{r, echo = FALSE}
w_plot <- ggplot(noshow, aes(x = Status, y = WaitingTimeDays, fill = Status))
w_plot + geom_boxplot() + ggtitle("Waiting time in days by Status") + xlab("Status group") + ylab("Waiting time in days")

```

It seems that scale of the waiting time is so twisted that it makes the interpretation quite hard.

```{r, echo = FALSE}
summary(noshow$WaitingTimeDays)
```

We need to divide the variable into intervals. Quantile points of the variable WaitingTimeDays are: 

```{r, echo = FALSE}
#making the intervals
bins <- quantile(noshow$WaitingTimeDays)

wt <- noshow$WaitingTimeDays
#create a categorical variable 'WaitingTime'
WaitingTimeIntervals <- cut(wt, breaks = bins, include.lowest = TRUE, label = c("short", "med_short", "med_long", "long"))

#look at the table of the new factor wti
wti_tab <-table(WaitingTimeIntervals)
wti_tab

# add the new categorical value to noshow
noshow <- data.frame(noshow, WaitingTimeIntervals)
#str(noshow)

```


```{r, echo = FALSE}
ggplot(noshow, aes(WaitingTimeIntervals, fill = Status)) + geom_bar(position = "fill") + ggtitle("Propotion of patients in status groups by waiting time")

```

It seems that the propotion of the no-shows is getting higher when the waiting time is growing.


##Diseases and habits as binary variables##

```{r, echo = FALSE}
dia_plot <- ggplot(noshow, aes(Diabetes, fill = Status)) + geom_bar(position = "fill")
alc_plot <- ggplot(noshow, aes(Alcoholism, fill = Status)) + geom_bar(position = "fill")
hip_plot <- ggplot(noshow, aes(HipTension, fill = Status)) + geom_bar(position = "fill")
tube_plot <- ggplot(noshow, aes(Tuberculosis, fill = Status)) + geom_bar(position = "fill")
grid.arrange(dia_plot, alc_plot, hip_plot, tube_plot, ncol = 2, top = "Deseases as binary variables")
```

It seems that patients who has deseases like diabetes or hip tensions that require regular visit in hospital tend to be more prone to show up on appointment. Patients who suffer from alcoholism or tuberculosis do not visit hospital on regular basis and maybe therefore they are more prone to have no-shows.


```{r, echo = FALSE}
hand_plot <- ggplot(noshow, aes(Handicap, fill = Status)) + geom_bar(position = "fill")
smo_plot <- ggplot(noshow, aes(Smokes, fill = Status)) + geom_bar(position = "fill")
sch_plot <- ggplot(noshow, aes(Scholarship, fill = Status)) + geom_bar(position = "fill")
sms_plot <- ggplot(noshow, aes(Sms_Reminder, fill = Status)) + geom_bar(position = "fill")
grid.arrange(hand_plot, smo_plot, sch_plot, sms_plot, ncol = 2, top = "Habits as binary variables")
```

First glimpse to habits shows that the is no remarkable differences between no-show and show up patients. It suprising that even sms reminder do not make difference among patients at this data set.


#Part two: Logistic regression model#

This is the second chapter of the final assignment. In this analysis we are using logistic regression model and trying predict patients who do not show up to appointment. In chapter one we reviewed differend variables against *Status* and find that others have greater variance than others.

##Explanatory variables##

After careful consideration we use the following variables in logistic regression model

|variable|description|
|---------|----------|
|Age|numeric|
|WaitingTimeIntervals|factor (1-4 levels)|
|Alcoholism|boolean (true/false)|
|Diabetes|boolean (true/false)|
|Scholarship|boolean (true/false)|

```{r, echo = FALSE}
model <- glm(Status ~ Age + WaitingTimeDays + Alcoholism + Diabetes + Scholarship, data = noshow, family = "binomial")
summary(model)
```

##Odds ratio##

```{r, echo = FALSE}
# compute odds ratios (OR)
OR <- coef(model) %>% exp
# compute confidence intervals (CI)
CI <- confint(model) %>% exp
#print out the odds ratios
cbind(OR, CI)
```

Odds ratio (OR) and the 95% confidence interval (CI) shows that those patients who have 


