#Chapter 1: Introduction to data

This is a analysis of EDA Medical Apppointment No-show data. Data is from [Kaggle](https://www.kaggle.com/joniarroba/noshowappointments). First we are going to have a good overview of the data and then we are going to predict show-ups and no-shows with some statistical modelling.

This analysis is the final assignment of the IODS course. You can find the [GitHub repository here.](https://github.com/mkkjrv/IODS-final). Author of this analysis is Mikko Järvi *mikko.jarvi(at)helsinki.fi*

```{r, include = FALSE}
library(ggplot2)
library(GGally)
library(lubridate)
```

##Overview of data##

After cleaning the the original dataset there is 15 variables and 300 000 observations. All the variables were characters in the original dataset. For the analysis the variables were converted as follows. Script for the data wrangling is [here.](https://github.com/mkkjrv/IODS-final/blob/master/data/create_noshow.R)

```{r, echo = F}
noshow <- read.csv("data/noshow.csv", header = TRUE)
noshow <- dplyr::select(noshow, -X)
noshow$AppointmentRegistration <- ymd_hms(noshow$AppointmentRegistration)
noshow$ApointmentData <- ymd_hms(noshow$ApointmentData)
str(noshow)
summary(noshow)
```



```{r, echo = FALSE}
#ggpairs(noshow, mapping = aes(col = Gender, alpha = 0.1), lower = list(combo = #wrap("facethist", bins = 20)))
``` 




