#Chapter 2: Analysis of the data

This is the second chapter of the final assignment. In this analysis we are using logistic regression model and trying predict patients who do not show up to appointment. In chapter one we reviewed differend variables against *Status* and find that others have greater variance than others.

##Explanatory variables##

After careful consideration we use the following variables in logistic regression model

|variable|description|
|---------|----------|
|Age|numeric|
|WaitingTimeIntervals|factor (1-4 levels)|
|Alcoholism|boolean (true/false)|
|Diabetes|boolean (true/false)|
|Scholarship|boolean (true/false)|

```{r, echo = FALSE}
model <- glm(Status ~ Age + WaitingTimeDays + Alcoholism + Diabetes + Scholarship, data = noshow, family = "binomial")
summary(model)
```

```{r, echo = FALSE}
# compute odds ratios (OR)
OR <- coef(model) %>% exp
# compute confidence intervals (CI)
CI <- confint(model) %>% exp
#print out the odds ratios
cbind(OR, CI)
```

Odds ratio (OR) and the 95% confidence interval (CI) shows that those patients who have 

